---
title: "Sentiment Analysis Project"
author: "Malayas, Pauchano, Madayag"
date: "2024-12-12"
output: pdf_document
---
 - BSIT-2A
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(tidyverse)
library(syuzhet)

tweetsDF <- read.csv("tweetsDF.csv")
full_df <- tweetsDF
clean <- full_df %>%
  mutate(text = str_to_lower(text),                     
         text = str_remove_all(text, "http\\S+"),       
         text = str_remove_all(text, "@\\w+"),          
         text = str_remove_all(text, "#\\w+"),          
         text = str_remove_all(text, "\\d+"),           
         text = str_remove_all(text, "[^\\w\\s]"),     
         text = str_squish(text),                      
         sentiment = get_sentiment(text, method = "bing")) 

clean <- clean %>%
  mutate(date = ymd_hms(created)) %>%  
  mutate(hour = hour(date))           
```

  - For the trend analysis, we examined the frequency of tweets posted every hour about the Itaewon incident to see how tweet activity evolved over time. We hoped to discover when key updates or news about the incident caused surges in user involvement by examining tweet counts at different hours. This data enables us to better understand the timing of public reactions and responses, as well as how information disseminated and when people were most active in discussing the incident.

```{r}

hourly_summary <- clean %>%
  group_by(day = as.Date(date), hour) %>%
  summarise(tweet_count = n(), .groups = "drop")
day1_tweets <- hourly_summary %>% filter(day == unique(day)[1])
day2_tweets <- hourly_summary %>% filter(day == unique(day)[2])
day3_tweets <- hourly_summary %>% filter(day == unique(day)[3])
```


```{r}

day1_graph <- ggplot(day1_tweets, aes(x = factor(hour), y = tweet_count, fill = tweet_count)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Tweet Frequency by Hour -", unique(day1_tweets$day)),
       x = "Hour of the Day",
       y = "Number of Tweets") +
  scale_fill_gradient(low = "green", high = "red") +
  scale_x_discrete(breaks = as.character(0:23)) +
  theme_minimal() 


day2_graph <- ggplot(day2_tweets, aes(x = factor(hour), y = tweet_count, fill = tweet_count)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Tweet Frequency by Hour -", unique(day2_tweets$day)),
       x = "Hour of the Day",
       y = "Number of Tweets") +
  scale_fill_gradient(low = "green", high = "red") +
  scale_x_discrete(breaks = as.character(0:23)) +
  theme_minimal() 

day3_graph <- ggplot(day3_tweets, aes(x = factor(hour), y = tweet_count, fill = tweet_count)) +
  geom_bar(stat = "identity") +
  labs(title = paste("Tweet Frequency by Hour -", unique(day3_tweets$day)),
       x = "Hour of the Day",
       y = "Number of Tweets") +
  scale_fill_gradient(low = "green", high = "red") +
  scale_x_discrete(breaks = as.character(0:23)) +
  theme_minimal()
```
Visualization and explanations of the graphs for Trend Analysis
```{r}
# 2022-10-28
day1_graph
```
  - The first graph shows very low tweet activity throughout the day, with no significant peaks or surges. This pattern indicates that the day was relatively uneventful in terms of the incident in focus. The absence of notable tweet volume suggests that discussions were likely unrelated to the tragedy, as it had not yet occurred or gained any attention.

```{r}
# 2022-10-29
day2_graph
```
  - The second graph indicates a significant rise in tweet activity starting around 3 PM, with a pronounced peak around 6 PM. This surge corresponds to the unfolding of the tragedy, as the incident likely occurred or was reported widely during this period. The increase in tweets shows how social media became a key platform for spreading information, expressing reactions, and coordinating responses as the tragedy unfolded.

```{r}
# 2022-10-30
day3_graph
```
  - The third graph displays a high volume of tweets during the early morning hours, peaking between midnight and 3 AM. Afterward, there is a steady decline in tweet activity throughout the rest of the day. The elevated tweet volume during the early hours likely reflects the aftermath of the Itaewon tragedy, as people continued sharing updates, reactions, and discussions in real time. The gradual decline suggests that as the day progressed, the urgency to tweet about the incident decreased, possibly due to the dissemination of official information or a shift in public focus.

END OF TREND ANALYSIS 



Preparing data for Sentiment Analysis

  - For the sentiment analysis, we focused on categorizing tweets into positive, neutral, and negative sentiments to observe emotional patterns throughout the day. This approach helps us understand how sentiments shift over time, providing valuable insights for responding to audience reactions.

```{r}
# Categorize sentiment into Positive, Neutral, and Negative by hour per day
sentiment_summary <- clean %>%
  mutate(sentiment_category = ifelse(sentiment > 0, "Positive",
                                      ifelse(sentiment < 0, "Negative", "Neutral"))) %>%
  group_by(day = as.Date(date), hour, sentiment_category) %>%
  summarise(count = n(), .groups = "drop")

# Filter each day's data for plotting graphs
day1_sentiments <- sentiment_summary %>% filter(day == unique(day)[1])
day2_sentiments <- sentiment_summary %>% filter(day == unique(day)[2])
day3_sentiments <- sentiment_summary %>% filter(day == unique(day)[3])
```

--- Codes for plotting sentiment analysis ---
```{r}
# Plot sentiment trends for the first day
sentiment_day1_plot <- ggplot(day1_sentiments, aes(x = factor(hour), y = count, fill = sentiment_category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = paste("Sentiment Distribution by Hour -", unique(day1_sentiments$day)),
       x = "Hour of the Day",
       y = "Number of Tweets") +
  scale_fill_manual(values = c("red", "yellow", "green")) +
  scale_x_discrete(breaks = as.character(0:23)) +
  theme_minimal()

# Plot sentiment trends for the second day
sentiment_day2_plot <- ggplot(day2_sentiments, aes(x = factor(hour), y = count, fill = sentiment_category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = paste("Sentiment Distribution by Hour -", unique(day2_sentiments$day)),
       x = "Hour of the Day",
       y = "Number of Tweets") +
  scale_fill_manual(values = c("red", "yellow", "green")) +
  scale_x_discrete(breaks = as.character(0:23)) +
  theme_minimal()

# Plot sentiment trends for the third day
sentiment_day3_plot <- ggplot(day3_sentiments, aes(x = factor(hour), y = count, fill = sentiment_category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = paste("Sentiment Distribution by Hour -", unique(day3_sentiments$day)),
       x = "Hour of the Day",
       y = "Number of Tweets") +
  scale_fill_manual(values = c("red", "yellow", "green")) +
  scale_x_discrete(breaks = as.character(0:23)) +
  theme_minimal()

```
